Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-23 paper/icml_drau/advup.tex

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
Still, this reparameterization does not, on its own, guarantee that $A$
correctly approximates $A^\pi_\deltat$ if
$Q_\Theta$ approximates $Q^\pi_\deltat$% \footnote{
% 	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
% 	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
% 	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is
% 	the same. \TODO{still unclear why we'd care about identifiability}
%}
.
Indeed, for any given pair $(V_\theta,\,A_\psi)$, the pair $(V_\theta(s) -
f(s),\,A_\psi(s,a)+f(s)/\deltat)$ (for an arbitrary $f$)
% \begin{align}
% 	\tilde{V}(s) &= V(s) - f(s),\qquad
% 	\tilde{A}(s, a) = A(s, a) + \frac{f(s)}{\deltat}
% \end{align}
yields the exact same function $Q_\Theta$. This new $A_\psi$ still defines %the same $Q$ and
the same ranking of actions, yet this phenomenon might cause
numerical problems or instability of $A_\psi$ when $\deltat\to 0$, and prevents direct
interpretation of the learned $A_\psi$.
% Consequently, $Q$ can correctly approximate
% $Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A_\psi$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A_\psi$ and $V_\theta$. This translates to
\begin{equation}
	A_\psi(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q_\Theta = Q^\pi_\deltat$, then $A_\psi =
A^\pi_\deltat$ and $V_\theta = V^\pi_\deltat$: indeed
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q_\Theta(s,a) - Q_\Theta(s, \pi(s))}{\deltat}
		    %\\ &
		    = A_\psi(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, instead of directly parameterizing $A_\psi$,
we define a parametric function $\bar{A}_\psi$ (typically a neural network),
and use $\bar{A}_\psi$ to define $A_\psi$ as
\begin{equation}
\label{eq:Aparam}
	A_\psi(s, a) \deq \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s)).
\end{equation}
% Rq Harsh
With this definition, $A_\psi$ directly verifies the normalization condition.

@@

To obtain a timestep-invariant RL algorithm, a timestep-invariant
exploration scheme is required. 
For continuous actions, \cite{ddpg} already introduced a time discretization invariant exploration scheme, which consists in adding an
\emph{Ornstein--Uhlenbeck} \cite{orn-uhl} (OU) process to actions. Formally, it is defined as
\begin{equation}
	\pi^\text{explore}(s^k_\deltat, z^k_\deltat) \deq \pi(s^k_\deltat) + z^k_\deltat
\end{equation}
with $z^k_\deltat$ the discretization of a continuous-time OU process,
\begin{equation}
	dz_t = - z_t \,\kappa\, dt + \sigma \,\dbt.
	\label{eq:orn_uhl}
\end{equation}
where $B_t$ is a brownian motion, $\kappa$ a stiffness parameter and
$\sigma$ a noise scaling parameter. The discretized trajectories converge
to non trivial continuous time trajectories, exhibiting a Brownian
behavior with a recall force towards $0$.

>>>>

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
Still, this reparameterization does not, on its own, guarantee that $A$
correctly approximates $A^\pi_\deltat$ if
$Q_\Theta$ approximates $Q^\pi_\deltat$% \footnote{
% 	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
% 	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
% 	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is
% 	the same. \TODO{still unclear why we'd care about identifiability}
%}
.
Indeed, for any given pair $(V_\theta,\,A_\psi)$, the pair $(V_\theta(s) -
f(s),\,A_\psi(s,a)+f(s)/\deltat)$ (for an arbitrary $f$)
% \begin{align}
% 	\tilde{V}(s) &= V(s) - f(s),\qquad
% 	\tilde{A}(s, a) = A(s, a) + \frac{f(s)}{\deltat}
% \end{align}
yields the exact same function $Q_\Theta$. This new $A_\psi$ still defines %the same $Q$ and
the same ranking of actions, yet this phenomenon might cause
numerical problems or instability of $A_\psi$ when $\deltat\to 0$, and prevents direct
interpretation of the learned $A_\psi$.
% Consequently, $Q$ can correctly approximate
% $Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A_\psi$, one must enforce the consistency equation
\begin{equation}
	V^\pi_\deltat(s) = Q^\pi_\deltat(s,
	\pi(s))
\end{equation}
on the approximate $A_\psi$ and $V_\theta$. This translates to
\begin{equation}
	A_\psi(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q_\Theta = Q^\pi_\deltat$, then $A_\psi =
A^\pi_\deltat$ and $V_\theta = V^\pi_\deltat$: indeed
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q_\Theta(s,a) - Q_\Theta(s, \pi(s))}{\deltat}
		    %\\ &
		    = A_\psi(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, instead of directly parameterizing $A_\psi$,
we define a parametric function $\bar{A}_\psi$ (typically a neural network),
and use $\bar{A}_\psi$ to define $A_\psi$ as
\begin{equation}
\label{eq:Aparam}
	A_\psi(s, a) \deq \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s)).
\end{equation}
% Rq Harsh
so that $A_\psi$ directly verifies the consistency condition.

@@

To obtain a timestep-invariant RL algorithm, a timestep-invariant
exploration scheme is required. 
For \emph{continuous} actions, \cite{ddpg} already introduced such a scheme, 
by adding an
\emph{Ornstein--Uhlenbeck} \cite{orn-uhl} (OU) process to actions. Formally, it is defined as
\begin{equation}
	\pi^\text{explore}(s_{k\deltat}, z_{k\deltat}) \deq
	\pi(s_{k\deltat}) + z_{k\deltat}
\end{equation}
with $z_{k\deltat}$ the discretization of a continuous-time OU process,
\begin{equation}
	dz_t = - z_t \,\kappa\, dt + \sigma \,\dbt.
	\label{eq:orn_uhl}
\end{equation}
where $B_t$ is a brownian motion, $\kappa$ a stiffness parameter and
$\sigma$ a noise scaling parameter. The discretized trajectories converge
to non trivial continuous time trajectories, exhibiting a Brownian
behavior with a recall force towards $0$.

### 2019-01-23 paper/icml_drau/icml_drau.tex
emacs
.

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2018}

@@

\begin{icmlauthorlist}
%\icmlauthor{Aeiau Zzzz}{equal,to}
%\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
%\icmlauthor{Cieua Vvvvv}{goo}
%\icmlauthor{Iaesut Saoeu}{ed}
%\icmlauthor{Fiuea Rrrr}{to}
%\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
%\icmlauthor{Aaoeu Iasoh}{goo}
%\icmlauthor{Buiui Eueu}{ed}
%\icmlauthor{Aeuia Zzzz}{ed}
%\icmlauthor{Bieea C.~Yyyy}{to,goo}
%\icmlauthor{Teoau Xxxx}{ed}
%\icmlauthor{Eee Pppp}{ed}
\end{icmlauthorlist}

%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
%\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
%
%\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

@@

\begin{abstract}
	Despite remarkable successes, \emph{Deep Reinforcement Learning}
	(DRL) is not robust to hyperparameterization, implementation
	details, or small environment changes (\citealt{drl_matter},
	\citealt{drl_matter_bis}).  Overcoming such sensitivity is key to
	making DRL applicable to real world problems.  In this paper,
	we identify
	sensitivity to \emph{time discretization} in near
	continuous-time environments as a critical factor; this covers, e.g., changing the number of
	frames per second, or the action frequency of the controller.
	Empirically, we find that
	Q-learning-based approaches such as
	\emph{Deep Q-learning}~\cite{dqn} and \emph{Deep Deterministic
	Policy Gradient}~\cite{ddpg} %are highly sensitive to variations of time discretization and
	collapse with small time steps. Formally, we prove that
	$Q$-learning does not exist in
	continuous time.
	We detail a principled way to build an
	off-policy RL algorithm that yields similar performances over a wide range
	of time
	discretizations, and confirm this robustness empirically. 
	% We further show that a simple reparameterization of the
	% Q-function as a sum of a state value term and a \emph{small}
	% action dependent advantage term yields an algorithm much more
	% resilient to variations of time discretization.\LB{Change last
	% sentence to: We introduce a more resilient algorithm to
	% variations of time discretization [We could say more, but still
	% less that what is already there]}
\end{abstract}

\input{intro.tex}

\input{relatedworks.tex}

\input{framework.tex}

\input{advup.tex}

\input{experiments.tex}

\input{discussions.tex}

\input{conclusion.tex}

\bibliography{icml_drau.bib}
\bibliographystyle{icml2018}
\end{document}

>>>>

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2019}

@@

\begin{icmlauthorlist}
	\icmlauthor{Corentin Tallec}{tau}
	\icmlauthor{L\'eonard Blier}{tau,fair}
	\icmlauthor{Yann Ollivier}{fair}
%\icmlauthor{Aeiau Zzzz}{equal,to}
%\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
%\icmlauthor{Cieua Vvvvv}{goo}
%\icmlauthor{Iaesut Saoeu}{ed}
%\icmlauthor{Fiuea Rrrr}{to}
%\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
%\icmlauthor{Aaoeu Iasoh}{goo}
%\icmlauthor{Buiui Eueu}{ed}
%\icmlauthor{Aeuia Zzzz}{ed}
%\icmlauthor{Bieea C.~Yyyy}{to,goo}
%\icmlauthor{Teoau Xxxx}{ed}
%\icmlauthor{Eee Pppp}{ed}
\end{icmlauthorlist}

\icmlaffiliation{tau}{TAckling the Underspecified, Universit\'e Paris Sud}
\icmlaffiliation{fair}{Facebook Artificial Intelligence Research}
%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
%
%\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

@@

\begin{abstract}
Despite remarkable successes, \emph{Deep Reinforcement Learning}
(DRL) is not robust to hyperparameterization, implementation
details, or small environment changes (\citealt{drl_matter},
\citealt{drl_matter_bis}).  Overcoming such sensitivity is key to
making DRL applicable to real world problems.  In this paper,
we identify
sensitivity to \emph{time discretization} in near
continuous-time environments as a critical factor; this covers, e.g., changing the number of
frames per second, or the action frequency of the controller.
Empirically, we find that
$Q$-learning-based approaches such as
\emph{Deep $Q$-learning}~\cite{dqn} and \emph{Deep Deterministic
Policy Gradient}~\cite{ddpg} %are highly sensitive to variations of time discretization and
collapse with small time steps. Formally, we prove that
$Q$-learning does not exist in
continuous time.
We detail a principled way to build an
off-policy RL algorithm that yields similar performances over a wide range
of time
discretizations, and confirm this robustness empirically. 
% We further show that a simple reparameterization of the
% Q-function as a sum of a state value term and a \emph{small}
% action dependent advantage term yields an algorithm much more
% resilient to variations of time discretization.\LB{Change last
% sentence to: We introduce a more resilient algorithm to
% variations of time discretization [We could say more, but still
% less that what is already there]}
\end{abstract}

\input{intro.tex}
\input{relatedworks.tex}
\input{framework.tex}
\input{advup.tex}
\input{experiments.tex}
\input{discussions.tex}
\input{conclusion.tex}
\pagebreak
\bibliography{icml_drau.bib}
\bibliographystyle{icml2019}
\end{document}
